Human capital mediates natural selection in contemporary humans

### 2022-05-04 why-natural-selection.Rmd
Final final final tweaks. Never say final.

res_wt_combined %>% 
      filter(Weights %in% c("Main", "None")) %>% 
      mutate(
        Weights  = recode(Weights, "Main" = "Weighted", "None" = "Unweighted"),
        term     = pretty_names(term),
        term     = fct_reorder(term, estimate, order_abs(1))
      ) %>% 
      standard_ggplot(fill_col = Weights, n_regs = n_regs, fill_direction = -1)

@@

```{r plot-cors-earnings-educ, fig.cap="Selection effects by correlations with earnings and educational attainment. Each point represents one polygenic score."}

@@

Because selection effects are concentrated in lower-income groups, they may also
increase inequality with respect to polygenic scores. For example, Figure
\@ref(fig:plot-mean-EA3-by-income) plots mean polygenic scores for educational
attainment (EA3) among children from households of different income groups. The
blue bars show the actual means, i.e. parents' mean polygenic score weighted by
number of children. The grey bars show the hypothetical means if all households
had equal numbers of children. Natural selection against genes associated with
educational attainment, which lowers the mean, is stronger at the bottom of the
income distribution, and this increases the differences between groups.
Overall, natural selection increases the correlation of polygenic scores with
income for `r n_increases` out of 33 polygenic scores, with a median percentage
increase of `r median_pct_change`% in the respondents' generation (Appendix
Table \@ref(tab:tbl-inequality-effects)).  If inequalities in polygenic scores
are important for understanding social structure and mobility
[@belsky2018genetic; @Rimfeld_2018; @harden2021genetic], then these increases
are substantive. Similarly, since many polygenic scores are predictive of
disease risk, they could potentially increase health inequalities. In general,
the evolutionary history of anatomically modern humans is related to disease
risk [@benton2021influence]; understanding the role of contemporary natural
selection may help researchers to map the genetic architecture of health
disparities.

@@

res_sex %>% 
      standard_ggplot(fill_col = sex, n_regs = n_regs, score_col = term,
                        conf_int = FALSE) +
      geom_linerange(data = res_sex_wide, inherit.aes = FALSE, 
                     color = "steelblue4", alpha = 1,
              aes(
                y      = term,
                x      = NULL,
                xmin   = min_est + 0.0005, 
                xmax   = max_est - 0.0005,
                linetype = p < 0.05/{{n_comps}}
              )) +
      scale_linetype_manual(values = c("TRUE" = 1, "FALSE" = 3))

@@

res_sibs_kids <-  bind_rows(
                    Siblings = res_sibs_parent_weights, 
                    Children = res_children_comparison,
                    .id = "Dep. var."
                  )

standard_ggplot(res_sibs_kids, `Dep. var.`, n_regs = length(score_names))
```

>>>>

res_wt_combined %>% 
      filter(Weights %in% c("Main", "None")) %>% 
      mutate(
        Weights  = recode(Weights, "Main" = "Weighted", "None" = "Unweighted"),
      ) %>% 
      standard_ggplot(fill_col = Weights, n_regs = n_regs, fill_direction = -1,
                        order_idx = 2)

@@

```{r plot-cors-earnings-educ, fig.cap="Selection effects by correlations with earnings and educational attainment. Each point represents one polygenic score. Selected scores are annotated."}

@@

Because selection effects are concentrated in lower-income groups, they may also
increase inequality with respect to polygenic scores. For example, Figure
\@ref(fig:plot-mean-EA3-by-income) plots mean polygenic scores for educational
attainment (EA3) among children from households of different income groups. The
red bars show the actual means, i.e. parents' mean polygenic score weighted by
number of children. The grey bars show the hypothetical means if all households
had equal numbers of children. Natural selection against genes associated with
educational attainment, which lowers the mean, is stronger at the bottom of the
income distribution, and this increases the differences between groups.
Overall, natural selection increases the correlation of polygenic scores with
income for `r n_increases` out of 33 polygenic scores, with a median percentage
increase of `r median_pct_change`% in the respondents' generation (Appendix
Table \@ref(tab:tbl-inequality-effects)).  If inequalities in polygenic scores
are important for understanding social structure and mobility
[@belsky2018genetic; @Rimfeld_2018; @harden2021genetic], then these increases
are substantive. Similarly, since many polygenic scores are predictive of
disease risk, they could potentially increase health inequalities. In general,
the evolutionary history of anatomically modern humans is related to disease
risk [@benton2021influence]; understanding the role of contemporary natural
selection may help researchers to map the genetic architecture of health
disparities.

@@

res_sex %>% 
      standard_ggplot(fill_col = sex, n_regs = n_regs, score_col = term,
                        conf_int = FALSE, fill_direction = -1) +
      geom_linerange(data = res_sex_wide, inherit.aes = FALSE, 
                     color = "steelblue4", alpha = 1,
              aes(
                y      = term,
                x      = NULL,
                xmin   = min_est + 0.0005, 
                xmax   = max_est - 0.0005,
                linetype = p < 0.05/{{n_comps}}
              )) +
      scale_linetype_manual(values = c("TRUE" = 1, "FALSE" = 3))

@@

res_sibs_kids <-  bind_rows(
                    Parents = res_sibs_parent_weights, 
                    Respondents = res_children_comparison,
                    .id = "Generation"
                  )

standard_ggplot(res_sibs_kids, Generation, n_regs = length(score_names),
                  fill_direction = -1)
```
